---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

#### Packages used

```{r}
#install.packages("tuneR")
#install.packages("ggplot2")
#install.packages("seewave")
#install.packages("signal")
library(tuneR)
library(ggplot2)
library(seewave)
library(signal)
```

### Task 1

1. Download the laugh.wav file [here](https://drive.google.com/file/d/1oju8aZPF8wCbQaWs6uXfochFBZ4U7xEH/view?usp=sharing)
2. Load it with the readWave function and plot its waveform.
3. Analyse the waveform, and load the file again, but only the last “hiccup”. Check the documentation of [readWave](https://www.rdocumentation.org/packages/tuneR/versions/1.4.5/topics/readWave) to identify the parameters you can use to load only a part of the file.

```{r}
# Code here

bomby_wav <- readWave("laugh.wav", from=(4.7), units="seconds")
plot(bomby_wav)

# duration_bomby <- duration(bomby_wav)
# bomby_wav
laugh_wav <- readWave("laugh.wav")
# duration_seconds <- duration(laugh_wav)
# print(duration_seconds)
# 5.041633
num_frames <- length(laugh_wav) # Measurement of the amplitude
max_amplitude <- max(audio@left) # Max amplitude of left channel
min_amplitude <- min(audio@left) # Min amplitude of left channel
file_size <- file.size("laugh.wav") # File size in bytes
sample_rate <- laugh_wav@samp.rate

print(paste0("Sample Rate = ", sample_rate, " Hz"))
```

### Task 2

Complete the `print_audio_metadata` function and print: 1) if the file is stereo or not, sample rate, bit depth, number of channels, duration in seconds, number of frames, min and max amplitude of each channel

```{r echo=TRUE}
print_audio_metadata <- function(file_name) {
  # Read the audio file
  audio <- readWave(file_name)

  # Extract properties
  is_stereo <- audio@stereo
  sample_rate <- audio@samp.rate
  bit_depth <- audio@bit
  duration_seconds <- duration(filename)
  num_frames <- length(laugh_wav)
  max_amplitude <- max(audio@left)
  min_amplitude <- min(audio@left)
  
  
  num_channels <- 2
  if (length(audio@right) == 0 || length(audio@left) == 0) {
    num_channels <- 1
  }
  
  # COMPLETE FUNCTION HERE
  
print(paste0("Stereo: True/False : ", num_channels))
print(paste0("Duraion (sec): ", duration_seconds)
print(paste0("Sample Rate: ", sample_rate ))
print(paste0(" Bit Depth: ", bit_depth)
print(paste0("Number of Channels: ", num_channels))     
print(paste0("Number of Frames: (Left)  ", num_frames))    
print(paste0("Max Amplitude: ", max_amplitude ))
print(paste0("Min Amplitude: ", min_amplitude ))
  

}

# Example usage:
print_audio_metadata("laugh.wav")

```

```{r}

audiofile_df <-("laugh.wav")

print_audio_metadata <- function(file_name) {
  # Read the audio file
  audio <- readWave(file_name)

  # Extract properties
  is_stereo <- audio@stereo
  sample_rate <- audio@samp.rate
  bit_depth <- audio@bit
  duration_seconds <- duration(audio)
  num_frames <- length(laugh_wav)
  max_amplitude <- max(audio@left)
  min_amplitude <- min(audio@left)
  
  
  num_channels <- 2
  if (length(audio@right) == 0 || length(audio@left) == 0) {
    num_channels <- 1
  }
  
print(paste0("Duration: ", duration(audio), " seconds"))
print(paste0("Sample Rate: ", sample_rate, "Hz\n"))
print(paste0("Bit Depth: ", bit_depth, "bits\n"))
print(paste0("Number of Channels: ", num_channels))
if (length(audio@left) > 0) {
    print(paste0("Number of Frames (Left): ", length(audio@left)))
    print(paste0("Min Amplitude (Left Channel): ", min(audio@left)))
    print(paste0("Max Amplitude (Left Channel): ", max(audio@left)))
}
if (length(audio@right) > 0) {
    print(paste0("Number of Frames (Right): ", audio@right))
    print(paste0("Min Amplitude (Right Channel): ", min(audio@right)))
    print(paste0("Max Amplitude (Right Channel): ", max(audio@right)))
}

}

getwd()
list.files()
print_audio_metadata(audiofile_df)


```



```{r}

library(tuneR)

print_audio_metadata <- function(file_name) {
  
  # Read the audio
  audio <- readWave(file_name)

  # Extract properties
  is_stereo        <- audio@stereo
  sample_rate      <- audio@samp.rate
  bit_depth        <- audio@bit
  duration_seconds <- duration(audio)
  
  # Frame counts
  num_frames_left  <- length(audio@left)
  num_frames_right <- if (is_stereo) length(audio@right) else 0
  
  # Amplitude values
  max_amplitude_left <- max(audio@left)
  min_amplitude_left <- min(audio@left)
  
  if (is_stereo) {
    max_amplitude_right <- max(audio@right)
    min_amplitude_right <- min(audio@right)
  }
  
  # Channels
  num_channels <- ifelse(is_stereo, 2, 1)
  
  # ---- PRINT ----

  print(paste0("File anme = ", file_name))
  print("---------------------------")
  print(paste0("Stereo: ", is_stereo))
  print(paste0("Duration: ", duration_seconds, " seconds"))
  print(paste0("Sample Rate: ", sample_rate, " Hz"))
  print(paste0("Bit Depth: ", bit_depth, " bits"))
  print(paste0("Number of Channels: ", num_channels))
  
  print(paste0("Frames (Left): ", num_frames_left))
  print(paste0("Min Amplitude (Left): ", min_amplitude_left))
  print(paste0("Max Amplitude (Left): ", max_amplitude_left))
  
  if (is_stereo) {
    print(paste0("Frames (Right): ", num_frames_right))
    print(paste0("Min Amplitude (Right): ", min_amplitude_right))
    print(paste0("Max Amplitude (Right): ", max_amplitude_right))
  }
}



print_audio_metadata("laugh.wav")

```




```{r}

library(tuneR)

print_audio_metadata <- function(file_name)
  
  # Read the audio
  audio <- readWave(file_name)

  # Extract properties
  is_stereo        <- audio@stereo
  sample_rate      <- audio@samp.rate
  bit_depth        <- audio@bit
  duration_seconds <- duration(audio)
  
  # Frame counts
  num_frames_left  <- length(audio@left)
  num_frames_right <- if (is_stereo) length(audio@right) else 0
  
  # Amplitude values
  max_amplitude_left <- max(audio@left)
  min_amplitude_left <- min(audio@left)
  
  if (is_stereo) {
    max_amplitude_right <- max(audio@right)
    min_amplitude_right <- min(audio@right)
  }
  
  # Channels
  num_channels <- ifelse(is_stereo, 2, 1)
  

print(paste0("Duration: ", duration(audio), " seconds"))
print(paste0("Sample Rate: ", sample_rate, "Hz\n"))
print(paste0("Bit Depth: ", bit_depth, "bits\n"))
print(paste0("Number of Channels: ", num_channels))
if (length(audio@left) > 0) {
    print(paste0("Number of Frames (Left): ", length(audio@left)))
    print(paste0("Min Amplitude (Left Channel): ", min(audio@left)))
    print(paste0("Max Amplitude (Left Channel): ", max(audio@left)))
}
if (length(audio@right) > 0) {
    print(paste0("Number of Frames (Right): ", audio@right))
    print(paste0("Min Amplitude (Right Channel): ", min(audio@right)))
    print(paste0("Max Amplitude (Right Channel): ", max(audio@right)))
    

    if (numchannels > 2) {
    print(paste0("Cnogratualations yu have more than two channels and have Quadriphonc, dolby or some multi channel audio"))
}
    
}

print_audio_metadata("790888__maciekkubera__garden-at-dawn-warsaw-suburbs-1.wav")

```


### Task 3

Some experimentation is usually necessary when applying filters. Play with the butter n and W filter values. Compare different results and see if you can better approximate the filtered signal to the original one.

```{r}

t <- seq(0, 1, len = 100)
sig <- sin(2 * pi * t)
noisySig <- sin(2 * pi * t) + 0.25 * rnorm(length(t))

butterFilter <- butter(3, 0.1)
recoveredSig <- filter(butterFilter, noisySig)
allSignals <- data.frame(t, sig, noisySig, recoveredSig)
ggplot(allSignals, aes(t)) + 
  geom_line(aes(y = sig, color = "Original")) + 
  geom_line(aes(y = noisySig, color = "Noisy")) + 
  geom_line(aes(y = recoveredSig, color = "Recovered")) + 
  labs(x = "Time", y = "Signal")

```

### Task 4

Download the bird song [here](https://drive.google.com/file/d/1plo2puYg8WNgSuMe3TT_TKDO60qLTVi5/view?usp=sharing).
Plot its spectrogram using `spectro`. Using the previous parameters, try to provide a clearer representation of the bird sounds.

```{r}
# Code here

```
