---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

#### Packages used

```{r}
#install.packages("tuneR")
#install.packages("ggplot2")
#install.packages("seewave")
#install.packages("signal")
library(tuneR)
library(ggplot2)
library(seewave)
library(signal)
```

### Task 1

1. Download the laugh.wav file [here](https://drive.google.com/file/d/1oju8aZPF8wCbQaWs6uXfochFBZ4U7xEH/view?usp=sharing)
2. Load it with the readWave function and plot its waveform.
3. Analyse the waveform, and load the file again, but only the last “hiccup”. Check the documentation of [readWave](https://www.rdocumentation.org/packages/tuneR/versions/1.4.5/topics/readWave) to identify the parameters you can use to load only a part of the file.

```{r}
# Code here

```

### Task 2

Complete the `print_audio_metadata` function and print: 1) if the file is stereo or not, sample rate, bit depth, number of channels, duration in seconds, number of frames, min and max amplitude of each channel

```{r echo=TRUE}
print_audio_metadata <- function(file_name) {
  # Read the audio file
  audio <- readWave(file_name)

  # Extract properties
  is_stereo <- audio@stereo
  sample_rate <- audio@samp.rate
  bit_depth <- audio@bit
  
  num_channels <- 2
  if (length(audio@right) == 0 || length(audio@left) == 0) {
    num_channels <- 1
  }
  
  # COMPLETE FUNCTION HERE
  

}

# Example usage:
print_audio_metadata("laugh.wav")

```

### Task 3

Some experimentation is usually necessary when applying filters. Play with the butter n and W filter values. Compare different results and see if you can better approximate the filtered signal to the original one.

```{r}

t <- seq(0, 1, len = 100)
sig <- sin(2 * pi * t)
noisySig <- sin(2 * pi * t) + 0.25 * rnorm(length(t))

butterFilter <- butter(3, 0.1)
recoveredSig <- filter(butterFilter, noisySig)
allSignals <- data.frame(t, sig, noisySig, recoveredSig)
ggplot(allSignals, aes(t)) + 
  geom_line(aes(y = sig, color = "Original")) + 
  geom_line(aes(y = noisySig, color = "Noisy")) + 
  geom_line(aes(y = recoveredSig, color = "Recovered")) + 
  labs(x = "Time", y = "Signal")

```

### Task 4

Download the bird song [here](https://drive.google.com/file/d/1plo2puYg8WNgSuMe3TT_TKDO60qLTVi5/view?usp=sharing).
Plot its spectrogram using `spectro`. Using the previous parameters, try to provide a clearer representation of the bird sounds.

```{r}
# Code here

```
