---
title: "R Notebook"
output: html_notebook
---

# Lab Week

## Summary Statistics

### Wine dataset

Import and parse the wine dataset from the UCI machine learning repository into R. You can browse the contents of the dataset at <https://archive.ics.uci.edu/static/public/109/wine.zip>. Make sure you add column names and the correct column type for each column.

1.  Calculate the mean value for Alcohol.

```{r}

```

2.  Calculate the median value for Malic Acid

```{r}

```

3.  Calculate the five major percentile values for Flavanoids

```{r}

```

4.  Calculate the 23rd Percentile for Proanthocyanins

```{r}

```

5.  Calculate the mean value for all columns (this can be done as one liner)

```{r}

```

For the questions below, you might want to use a for loop to work through each column. You can print the results to the console.

6.  Calculate which column displays the greatest and least standard deviation

```{r}

```

7.  Using the standard deviation calculation, work out each column’s variance (do not use the built-in R function)

```{r}

```

8.  For each column, calculate the interquartile range

```{r}

```

9.  For each column, calculate whether the distribution is left or right skewed

```{r}

```

### Airquality dataset

Load the airquality dataset which is available with the baseR package. (You can use the R data function to load built-in datasets).

```{r}
data <- airquality
```

1.  Calculate the total number of NAs in the Solar.R column.

```{r}

```

2.  Calculate the total number of NAs in row 5

```{r}

```

3.  Calculate the mean of the Ozone column (the answer is not NA!).

```{r}

```

4.  Calculate the percentage of rows which contain at least 1 NA

```{r}

```

5.  Calculate the total number of NAs for each column

```{r}

```

## Graphs

Load the USArrests dataset from baseR.

```{r}
data <- USArrests
```

1.  Choose an appropriate visualisation (or visualisations) to determine whether UrbanPop or Assault are a better predictor of murder

```{r}

```

2.  Load the msleep dataset from baseR. Choose an appropriate visualisation to give a good summary of the spread of sleep patterns based on diet (the column vore). Plot this spread for each category of animal.

```{r}

```

3.  Load the titanic dataset. This can be installed using `install.packages("titanic")` and loaded with the `library()` function. After that, the dataset you want is called `titanic_train`. Choose an appropriate visualisation to show the distribution of ages on the titanic dataset.

```{r}

```

4.  Using the titanic dataset show how the likelihood of survival depends on gender and on class.

```{r}

```

5.  Using the diamonds dataset, investigate the relationship between carat and price. Use two geoms on a single plot. Is it linear?

```{r}

```

6.  Repeat the previous visualisation but show one graph for each different type of cut.

```{r}

```
