---
title: "R Notebook"
output: html_notebook
---

# Lab Week

## Summary Statistics

### Wine dataset

Import and parse the wine dataset from the UCI machine learning repository into R. You can browse the contents of the dataset at <https://archive.ics.uci.edu/static/public/109/wine.zip>. Make sure you add column names and the correct column type for each column.

1.  Calculate the mean value for Alcohol.

```{r}
# Download the wine.zip file from UCI
url <- 'https://archive.ics.uci.edu/static/public/109/wine.zip' # define the url
download.file(url, destfile = 'wine.zip') # downlaods the zip
unzip('wine.zip', list = T) # just has a look at whats inside it
unzip('wine.zip', exdir = './wine') # extracts to folder /wine directory
list.files('./wine') # lissts whats been extracted an dint he new folder


```
```{r}
# import the data intoa  dataframe or tibble
# When you use functions like read_csv() or read_delim(), the col_names argument tells R whether the first row of your data file contains column headers (names) or actual data.
# if you use col_names = TRUE , R assumes the first row of your file contains column names, not data
library(readr)
wine_tbl <- read_delim('./wine/wine.data', delim=',', col_names = FALSE)
spec(wine_tbl)


```

1s. check dataset type tibble or dataframe
```{r}

class(wine_tbl)
library(tibble)
is_tibble(wine_tbl)

# tibble  >>> into >>? dataframe
wine_tbl <- as.data.frame(wine_tbl)
class(wine_tbl)
is_tibble(wine_tbl)


# dataframe >>> into a >>>> tibble
wine_tbl <- as_tibble(wine_df)
class(wine_tbl)
is_tibble(wine_tbl)

```

2.  Calculate the median value for Malic Acid

```{r}

# When you use functions like read_csv() or read_delim(), the col_names argument tells R whether the first row of your data file contains column headers (names) or actual data.
# if you use col_names = TRUE , R assumes the first row of your file contains column names, not data
library(readr)
wine_tbl <- read_delim('./wine/wine.data', delim=',', col_names = FALSE)
spec(wine_tbl)

# assign proper column names from the wine.names file for the columns
colnames(wine_tbl) <- c(
  "Class",
  "Alcohol",
  "Malic_Acid",
  "Ash",
  "Alcalinity_of_Ash",
  "Magnesium",
  "Total_Phenols",
  "Flavanoids",
  "Nonflavanoid_Phenols",
  "Proanthocyanins",
  "Color_Intensity",
  "Hue",
  "OD280_OD315",
  "Proline"
)

median_malic <- median(wine_tbl$Malic_Acid)
paste (" The median valeu for mailc acid is ", median_malic)

```

```{r}

summary(wine_tbl)

```



3.  Calculate the five major percentile values for Flavanoids

```{r}

# only workls on single numeric column
# fivenum(wine_tbl)

# to ONLY APLY IT TO COLI,NS THATARE NUMERIC
sapply(wine_tbl[, sapply(wine_tbl, is.numeric)], fivenum)

```

4.  Calculate the 23rd Percentile for Proanthocyanins

```{r}

names(wine_tbl)
percntl_23 <- quantile(wine_tbl$Proanthocyanins, .23)

print(paste0("The 23rd percentile of Proanthocyanins is ", percntl_23))

```

5.  Calculate the mean value for all columns (this can be done as one liner)

```{r}

# lapply() is an R function that applies a function to each element of a list (or vector) and returns a list. Think of it as a “loop in one line.”
# sapply() is very similar to lapply() but with a key difference: it tries to simplify the result.

sapply(wine_tbl, function(x) mean(x, na.rm = TRUE))

# 🔍 Explanation:
#  - sapply() applies a function to each column of a data frame or tibble.
#  - mean(x, na.rm = TRUE) computes the mean for each column while ignoring NA values.
#  - It will return a named numeric vector showing the mean for each column.
# -  function(x) ? is a function without a name — defined on the fly and used inside another function (like sapply()). “For each element (here, each column) in wine_tbl, call it x, and then do something with it. ie.. get the mean .. 3 parameters in the brackets”
# function(x) defines an inline (anonymous) function.
# x represents one element from the data (in this case, each column).
# It’s commonly used in apply, lapply, sapply, map, and tidyverse functions.


# restrict it to numeric colmns only - "is.mnumeric"
# sapply automaticalyl returns a vector instead of a list so mean_list is a vector...
mean_list <- sapply(wine_tbl[sapply(wine_tbl, is.numeric)], mean, na.rm = TRUE)
print(mean_list)

median_list <- sapply(wine_tbl[sapply(wine_tbl, is.numeric)], median, na.rm = TRUE)
print(median_list)

```

For the questions below, you might want to use a for loop to work through each column. You can print the results to the console.

6.  Calculate which column displays the greatest and least standard deviation

```{r}

# sapply(wine_tbl, function(x) sd(x, na.rm = TRUE))

sd_list <-sapply(wine_tbl[sapply(wine_tbl, is.numeric)], sd, na.rm = TRUE)

# lastg element is not  -1 (-1 just returns all elements except frist element)
# to ge las element of a list have a few options
# sd_list <- c(0.124, 0.229, 0.274)
# sd_list[length(sd_list)]   # 0.274
# tail(sd_list, 1)           # 0.274
# sd_list[-1]                # 0.229 0.274 (drops the first element)

sd_list <- round(sort(sd_list), 3)
print(sd_list)

print(paste(" the min stanard deviatin avlue is : ", sd_list[1]))
print(paste(" the max stanard deviatin avlue is : ", tail(sd_list, 1)))
print(paste(" different  =  is : ", tail(sd_list, 1) - sd_list[1]))


```

7.  Using the standard deviation calculation, work out each column’s variance (do not use the built-in R function)

```{r}
# For Loop structure is different to python...

var_list <- sd_list^2
print(var_list)

for (i in var_list) {
  print(paste("variance is ", i))
}

```

8.  For each column, calculate the interquartile range

```{r}

iqr_list <-sapply(wine_tbl[sapply(wine_tbl, is.numeric)], IQR, na.rm = TRUE)

# sort in ascending order
iqr_list <- sort(iqr_list)

# pritn out column names also, unline earlier for loops which just took values..
for (name in names(iqr_list)) {
  print(paste(name, "IQR value is  :", iqr_list[name], "\n"))
}


```

9.  For each column, calculate whether the distribution is left or right skewed

```{r}

# ifelse(condition, value_if_true, value_if_false)
# is a vectorized if statement in R.
# It checks a condition for each element of a vector.
# Returns one of two values depending on whether that condition is TRUE or FALSE.

skew_direction <- ifelse(mean_list > median_list, "Right-skewed",
                    ifelse(mean_list < median_list, "Left-skewed",
                           "Approximately symmetric"))

# Creates a data frame — basically a table — combining several vectors that are the same length.
# So you’re building a summary table that has:
# Variable names,
# Their mean,
# Their median,
# Their skew direction (left/right/symmetric).


skew_summary <- data.frame(
  Variable = names(mean_list),
  Mean = round(mean_list, 3),
  Median = round(median_list, 3),
  Direction = skew_direction,
  row.names = NULL
)

# or create a tibble instead of dataframe above (nicer visualisatuin apprarently)
library(tibble)
skew_summary <- tibble(
  Variable = names(mean_list),
  Mean = round(mean_list, 3),
  Median = round(median_list, 3),
  Direction = skew_direction
)


print(skew_summary)

#optional visualisation
barplot(mean_list - median_list,
        main = "Mean − Median Difference (Skewness Direction)",
        ylab = "Mean − Median",
        col = ifelse(mean_list > median_list, "tomato", "steelblue"))
abline(h = 0, lty = 2)



```

### Airquality dataset

Load the airquality dataset which is available with the baseR package. (You can use the R data function to load built-in datasets).

```{r}
data <- airquality
str(data)
```

```{r}
str(data)
```

1.  Calculate the total number of NAs in the Solar.R column.

```{r}

# number of missing from solar column
sum(is.na(airquality$Solar.R))


# ask R
# allmissing <- is.na(data)
# print(allmissing)
# 
# # # get a count for eacjh column
# counts <- colSums(allMissing)
# print(counts)
# # only the columnsthat have missing NAs
# print(paste(" percentage NAs is ", (round(counts/nrow(data)*100, 2))))


# # check for missing values
# allMissing <- is.na(data)
# 
# # count NAs per column
# counts <- colSums(allMissing)
# 
# # calculate percentage of missing values per column
# percent_NA <- round(counts / nrow(data) * 100, 2)
# 
# # print results
# print(paste("Percentage of NAs is:", percent_NA))

```

2.  Calculate the total number of NAs in row 5

```{r}
sum(is.na(airquality[5, ]))


```

3.  Calculate the mean of the Ozone column (the answer is not NA!).

```{r}
mean_list <- sapply(data[sapply(data, is.numeric)], mean, na.rm = TRUE)
print(mean_list['Ozone'])
print(mean_list[1])

```

4.  Calculate the percentage of rows which contain at least 1 NA

```{r}

# check which rows have any NA
rows_with_NA <- rowSums(is.na(airquality)) > 0

# count how many rows have at least one NA
num_rows_with_NA <- sum(rows_with_NA)

# calculate percentage
percent_rows_with_NA <- (num_rows_with_NA / nrow(airquality)) * 100

# print results
cat("Number of rows with at least one NA:", num_rows_with_NA, "\n")
cat("Percentage of rows with at least one NA:", round(percent_rows_with_NA, 2), "%\n")



```

5.  Calculate the total number of NAs for each column

```{r}
# check which rows have any NA
cols_with_NA <- colSums(is.na(airquality)) 

print(cols_with_NA)

# 🧠 Explanation
# Part	What it does
# is.na(airquality)	Returns a logical matrix (TRUE for NAs, FALSE otherwise)
# colSums(...)	Adds up TRUEs column-wise (TRUE counts as 1, FALSE as 0)


```

## Graphs

Load the USArrests dataset from baseR.

```{r}
# can import as tibble using as_tibbel if we want
data_usarr <- as_tibble(USArrests)

summary(data_usarr)
class(data_usarr)
head(data_usarr)

# wine_tbl <- as_tibble(wine_df)
# class(wine_tbl)
# is_tibble(wine_tbl)
```

1.  Choose an appropriate visualisation (or visualisations) to determine whether UrbanPop or Assault are a better predictor of murder

```{r}
# scatterplto will give an idea of correlation bteween 2 variables
# 🧠 Step 1: Understanding your variables
# 
# In USArrests:
# # Variable	Meaning
# Murder	Murder arrests (per 100,000 residents)
# Assault	Assault arrests (per 100,000 residents)
# UrbanPop	Percent urban population
# Rape	Rape arrests (per 100,000 residents)
# 
# Goal:
# ➡️ Determine whether Assault or UrbanPop is more strongly related to Murder.
# ✅ Step 2: Choose appropriate visualizations
# Since all variables are numeric, you can use:
# ⃣ 1  Scatter plots
# # To visualize relationships between two numeric variables.
# # Add a regression line to see the strength and direction of the relationship.
# # 2️⃣ Correlation coefficients
# # To numerically compare the strength of relationships.

library(ggplot2)

# Scatter plot: Murder vs Assault
ggplot(data_usarr, aes(x = Assault, y = Murder)) +
  geom_point(color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "darkred") +
  labs(title = "Murder vs Assault",
       x = "Assault Arrests per 100,000",
       y = "Murder Arrests per 100,000")

# Scatter plot: Murder vs UrbanPop
ggplot(data_usarr, aes(x = UrbanPop, y = Murder)) +
  geom_point(color = "seagreen") +
  geom_smooth(method = "lm", se = FALSE, color = "darkred") +
  labs(title = "Murder vs UrbanPop",
       x = "Urban Population (%)",
       y = "Murder Arrests per 100,000")


cor(data_usarr$Murder, data_usarr$Assault)
cor(data_usarr$Murder, data_usarr$UrbanPop)

# 📊 Example output (actual results from USArrests)
# Variable	Correlation with Murder
# Assault	0.80 (strong positive)
# UrbanPop	0.07 (weak positive)
# 🧩 Step 5: Interpretation
# 
# The Murder vs Assault scatter plot will show a strong upward trend.
# 
# The Murder vs UrbanPop scatter plot will look scattered with no clear pattern.
# 
# The correlation confirms this numerically.
# 
# ✅ Conclusion:
# 
# Assault is a much better predictor of Murder than UrbanPop.

```

2.  Load the msleep dataset from baseR. Choose an appropriate visualisation to give a good summary of the spread of sleep patterns based on diet (the column vore). Plot this spread for each category of animal.

```{r}
library(tidyverse)
library(readr)
data_sleep <- as_tibble(msleep)
summary(data_sleep)
head(data_sleep)
names(data_sleep)
class(data_sleep)
# 
# 🟢 Best choice: Boxplot
# # A boxplot is ideal because it:
# # Shows distribution & spread (median, IQR, outliers)
# # Compares multiple categories
# # Handles numeric vs categorical relationships perfectly

ggplot(data_sleep, aes(x = vore, y = sleep_total, fill = vore)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Total Sleep Time by Diet Type",
    x = "Diet Category (vore)",
    y = "Total Sleep (hours per day)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# 🧠 Step 4: Interpretation
# # From this plot, you can easily see:
# # Which diet group sleeps the most or least
# # How variable sleep times are within each group
# # Whether distributions overlap or differ significantly
# # Typically (from this dataset):
# # Herbivores tend to sleep less
# # Carnivores and insectivores sleep more
# # Omnivores are in between

# improve by adding data points - 1️⃣ Add individual data points (to see spread better)
ggplot(msleep, aes(x = vore, y = sleep_total, fill = vore)) +
  geom_boxplot(alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.6, color = "black") +
  labs(
    title = "Spread of Total Sleep Time by Diet Type",
    x = "Diet Category",
    y = "Total Sleep (hours)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# 2️⃣ Use a violin plot for a smooth shape view

ggplot(msleep, aes(x = vore, y = sleep_total, fill = vore)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, fill = "white") +
  labs(
    title = "Spread of Total Sleep Time by Diet Type",
    x = "Diet Category",
    y = "Total Sleep (hours)"
  ) +
  theme_minimal()


```

3.  Load the titanic dataset. This can be installed using `install.packages("titanic")` and loaded with the `library()` function. After that, the dataset you want is called `titanic_train`. Choose an appropriate visualisation to show the distribution of ages on the titanic dataset.

```{r}
#install.packages("titanic")   # Run this once
#library(titanic)
data_tit <- (titanic_train)

# summary(data_tit)
# head(data_tit)
# names(data_tit)
# sum(is.na(data_tit$Age))
# class(data_tit)


# library(ggplot2)

# Option 1 - hiOption 1 - stogram
# What it shows:
# # Frequency of passengers in each age range (e.g., 0–5, 6–10, …).
# # You’ll see a peak around ages 20–35, meaning most passengers were young adults.
# Some passengers are missing ages (NA values).
ggplot(data_tit, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", na.rm = TRUE) +
  labs(
    title = "Distribution of Passenger Ages on the Titanic",
    x = "Age (years)",
    y = "Number of Passengers"
  ) +
  theme_minimal()

# Option 2 - Density PLot
# What it shows:
# # Smooth curve showing the probability density of ages.
# # Easier to compare if you later group by gender or survival.
ggplot(data_tit, aes(x = Age)) +
  geom_density(fill = "lightgreen", alpha = 0.6, na.rm = TRUE) +
  labs(
    title = "Density of Passenger Ages on the Titanic",
    x = "Age (years)",
    y = "Density"
  ) +
  theme_minimal()
# Option 3 - : Boxplot – shows spread, median, and outliers
# What it shows:
# Median age, interquartile range (IQR), and any extreme ages (children or elderly).
ggplot(data_tit, aes(y = Age)) +
  geom_boxplot(fill = "orange", outlier.color = "red", na.rm = TRUE) +
  labs(
    title = "Spread of Passenger Ages on the Titanic",
    y = "Age (years)"
  ) +
  theme_minimal()

# option 4 - Compare by Group i.e. first oen here is sex male female
  
  ggplot(data_tit, aes(x = Sex, y = Age, fill = Sex)) +
  geom_boxplot(na.rm = TRUE) +
  labs(
    title = "Distribution of Passenger Ages by Sex",
    x = "Sex",
    y = "Age (years)"
  ) +
  theme_minimal()

# or OOption 4 A - order by group  - second one here is survived factor yes or no 
ggplot(data_tit, aes(x = factor(Survived), y = Age, fill = factor(Survived))) +
  geom_boxplot(na.rm = TRUE) +
  labs(
    title = "Distribution of Passenger Ages by Survival Status",
    x = "Survived (0 = No, 1 = Yes)",
    y = "Age (years)"
  ) +
  theme_minimal()

# ✅ Step 4: Summary of the best choice
# Goal	Visualization	Why
# Show overall distribution of a numeric variable	Histogram	Most intuitive for spread & frequency
# Show smooth shape of distribution	Density plot	Highlights modes and skewness
# Show median, spread & outliers	Boxplot	Compact summary of variability

ggplot(data_tit, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black", na.rm = TRUE) +
  labs(
    title = "Distribution of Passenger Ages on the Titanic",
    x = "Age (years)",
    y = "Number of Passengers"
  ) +
  theme_minimal()

```

4.  Using the titanic dataset show how the likelihood of survival depends on gender and on class.

```{r}
# aggregate data
aggregate(Survived ~ Sex + Pclass, data = data_tit, mean)


# ✅ So survival was: 9from thetable produced from above
# # Much higher for females (especially in 1st & 2nd class)
# # Much lower for males, especially 3rd class passengers

# 📊 Step 3: Visualize with ggplot2
# Let’s show this clearly.
# Option 1: Bar plot of survival rates by gender and class
# > Interpretation of plto below:
# The y-axis shows proportion survived (not counts).
# Each bar adds up to 100%.
# You’ll see that female passengers, especially in 1st and 2nd class, had the highest survival rates.

ggplot(data_tit, aes(x = Sex, fill = factor(Survived))) +
  geom_bar(position = "fill") +
  facet_wrap(~ Pclass) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Likelihood of Survival by Gender and Passenger Class",
    x = "Gender",
    y = "Survival Rate (percentage)",
    fill = "Survived"
  ) +
  theme_minimal()

# Option 2: Grouped bar plot of mean survival
# 
# This one shows exact proportions per group.
# Interpretation:
# Bars show probability of survival.
# You can directly compare males vs females within each class.
library(dplyr)

survival_summary <- data_tit %>%
  group_by(Sex, Pclass) %>%
  summarise(SurvivalRate = mean(Survived, na.rm = TRUE))

# Plot
ggplot(survival_summary, aes(x = factor(Pclass), y = SurvivalRate, fill = Sex)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Survival Rate by Gender and Passenger Class",
    x = "Passenger Class",
    y = "Survival Rate",
    fill = "Gender"
  ) +
  theme_minimal()


# 🧩 Step 4: Key takeaway
# Group	Survival Rate (approx.)	Interpretation
# Female, 1st Class	~97%	Almost all survived
# Female, 2nd Class	~92%	Very high survival
# Female, 3rd Class	~50%	Half survived
# Male, 1st Class	~37%	Some survived
# Male, 2nd Class	~15%	Low survival
# Male, 3rd Class	~13%	Very low survival
# ✅ Step 5: Summary of visualization options
# Goal	Visualization	Best for Compare survival proportions visually	Stacked proportional bar chart (position="fill")	Easiest to interpret
# Show numeric survival rates by group	Grouped bar plot	Cleaner, exact values
# Summarize data numerically	aggregate() or dplyr::summarise()	Quick overview


```

5.  Using the diamonds dataset, investigate the relationship between carat and price. Use two geoms on a single plot. Is it linear?

```{r}

data_dmd <- as_tibble(diamonds)
summary(data_dmd)
head(data_dmd)
names(data_dmd)
class(data_dmd)
glimpse(data_dmd)

# ✅ Step 2: Create a scatter plot
## A scatter plot is ideal for showing the relationship between two numeric variables.
# # We’ll add two geoms:
# # geom_point() → raw data points
# # geom_smooth() → fitted trend line (linear or otherwise)

# 
# 🔍 Step 3: Interpretation
# You’ll notice:
# There’s a strong positive relationship: as carat increases, price increases.
# However, it’s not perfectly linear — the price increases faster than linearly for larger diamonds (a curved or exponential pattern).
# So, while a straight line (method = "lm") captures the general trend, it doesn’t fit the curvature well.

ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.3, color = "steelblue") +   # scatter plot
  geom_smooth(method = "lm", color = "red", se = FALSE) +  # linear trend line
  labs(
    title = "Relationship between Diamond Carat and Price",
    x = "Carat",
    y = "Price (USD)"
  ) +
  theme_minimal()

# 🟣 Step 4: Try a non-linear smoother for comparison
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.3, color = "skyblue") +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  labs(
    title = "Non-linear Fit: Carat vs Price",
    x = "Carat",
    y = "Price (USD)"
  ) +
  theme_minimal()

# 🧮 Optional: Quantify linearity
# You can test linearity with a model:
# model <- lm(price ~ carat, data = diamonds)
# summary(model)


# 📊 Step 5: (Optional) Use log transformation to check proportionality
# If you transform both variables using log, the relationship becomes closer to linear:

ggplot(diamonds, aes(x = log(carat), y = log(price))) +
  geom_point(alpha = 0.3, color = "darkgreen") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Log-Transformed Relationship between Carat and Price",
    x = "log(Carat)",
    y = "log(Price)"
  ) +
  theme_minimal()

# !!!!  ✅ The log–log plot shows a more linear relationship, suggesting that price grows exponentially with carat.!!!!!!
# 
# ✅ Final Summary
# Visualization	Geoms Used	What it shows
# Scatter + Linear fit	geom_point() + geom_smooth(method="lm")	Strong positive but not perfectly linear
# Scatter + LOESS smooth	geom_point() + geom_smooth(method="loess")	Captures curved, exponential-like trend
# Log–log plot	same geoms but with log() axes	Relationship becomes approximately linear

# ✍️ Interpretation:
# 
# There is a strong positive relationship between carat and price,
# but it is non-linear — the price increases more sharply as carat grows.
# In other words, larger diamonds cost disproportionately more.

ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.3, color = "skyblue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Carat vs Price: Relationship in the Diamonds Dataset",
    x = "Carat",
    y = "Price (USD)"
  ) +
  theme_minimal()


```

6.  Repeat the previous visualisation but show one graph for each different type of cut. - Facetg by cut

```{r}
# 🧠 Step 1: Recall the previous visualization
# # We used:
# # geom_point() + geom_smooth(method = "lm")
# # to show carat vs. price with both raw data and a trend line.
# # Now we’ll use faceting in ggplot2 to create one panel (graph) per cut type.

# ✅ Step 2: Code — facet by cut

ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.3, color = "skyblue") +         # scatter plot
  geom_smooth(method = "lm", color = "red", se = FALSE) +  # linear fit
  facet_wrap(~ cut) +                                  # one plot per cut type
  labs(
    title = "Option 2 - Relationship Between Carat and Price by Diamond Cut",
    x = "Carat",
    y = "Price (USD)"
  ) +
  theme_minimal()
# 
# 💡 Step 3: What this does
## facet_wrap(~ cut) creates separate subplots — one for each diamond cut.
#  Each subplot shows its own carat–price relationship.
#  You can visually compare:
# # Whether the slope differs by cut (i.e., how quickly price rises with carat)
# # Whether the spread of prices differs across quality grades

# 🔍 Step 4: Interpretation
# 
# When you run this, you’ll typically see:
# 
# Cut Type	Pattern Observed
# Fair	More variability; lower prices for a given carat
# Good / Very Good	Slightly higher prices and more consistent
# Premium / Ideal	Steeper relationship — price increases faster with carat size
# 
# So, higher-quality cuts tend to have a stronger price premium per carat — prices rise more sharply for Ideal and Premium cuts.
# 


# 🟢 Step 5: Optional — Use LOESS for non-linear trend
# 
# If you want to capture the curved (non-linear) relationship better:

ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.3, color = "skyblue") +
  geom_smooth(method = "loess", color = "darkred", se = FALSE) +
  facet_wrap(~ cut) +
  labs(
    title = "Option 5 - Non-linear Relationship Between Carat and Price by Cut",
    x = "Carat",
    y = "Price (USD)"
  ) +
  theme_minimal()

# ✅ Step 6: Optional Enhancement — color by cut (single plot version)
# 
# If you’d rather overlay all cuts in one plot for comparison:

ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "option 6 - Carat vs Price by Cut Type",
    x = "Carat",
    y = "Price (USD)"
  ) +
  theme_minimal()

# ✅ Final Answer (recommended version)
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 0.3, color = "skyblue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  facet_wrap(~ cut) +
  labs(
    title = "Rec final answer - Relationship Between Carat and Price by Diamond Cut",
    x = "Carat",
    y = "Price (USD)"
  ) +
  theme_minimal()


# ✍️ Interpretation:
# 
# Across all cuts, price increases with carat, but the relationship is non-linear (steeper for higher carats).
# Diamonds with higher cut quality (Ideal, Premium) tend to have higher prices at the same carat weight, and a steeper price–carat slope, reflecting quality premiums.


```
