---
title: "R Notebook"
output: html_notebook
---

# Visualizing Data

This notebook can be used to follow along with the practical examples from part 3 of this week's lecture. Before we begin make sure you've installed ggplot2. In these examples we'll be working with the *mpg* dataset. The mpg dataset deals with different makes and model of cars. The **hwy** column gives *highway miles per gallon*. This dataset is used to determine the factors that contribute to a car's miles per gallon

```{r}
# Uncomment the line below to install ggplot2 (if you haven't already).
install.packages('ggplot2')
library(ggplot2)

# The mpg dataset is available by default in R, it has the variable name `mpg`
nrow(mpg)
ncol(mpg)
head(mpg)
tail(mpg)

# count the number of distinct car models in this dataset
length(levels(as.factor(mpg$model)))
```

```{r}
# library(ggplot2)
# # ggplot(mpg, aes(x = displ, y = hwy)) + geom_point()
# 
# # colour
# ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(colour = 'red')
# 
# # size
# ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(size = 0.25)
# 
# 
# # transparency
# ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(alpha = 0.1)
# 
# # shape
# ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(shape = 4)


# move colour to aesthetic - 
# # colour
# ggplot(data = mpg, aes(x = displ, y = hwy, colour = class)) + geom_point()


# colour
ggplot(data = mpg, aes(x = displ, y = hwy, size = cty, colour = class)) + geom_point()

```

## Exploring qplot aesthetics

See the lecture notes for details. Plot the *hwy \~ displ* scatterplot above changing each of the 4 aesthetics

```{r}

```

Now setting `class` as the colour attribute

```{r}

```

## Mapping Aesthetics to Values

Use the colour aesthetic to represent each car's *class*

```{r}
# colour
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(colour = 'red')

# size
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(size = '0.25')


# transparency
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(alpha = '0.1')

# shape
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point(shape = '4')

```

## Faceting

Use faceting to plot the data in a grid of 1 column and 3 rows. The *drv* feature should determine which row an observation is plotted in

```{r}
ggplot(data = mpg, aes(x = displ, y = hwy)) + 
geom_point() + 
facet_grid(drv ~ . )

```

Now do the same, but rather than rows, have the facets in 3 columns

```{r}

ggplot(data = mpg, aes(x = displ, y = hwy)) + 
geom_point() + 
facet_grid(. ~ drv )


```

## Task 1 - Try each of the 3 examples of faceting

See lecture notes for examples to try

```{r}

# column displ vs hwy -- what is cyl?
# ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point() + facet_grid(. ~ cyl )

#ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point() + facet_grid(drv ~ cyl )
# 
ggplot(data = mpg, aes(x = displ, y = hwy)) + geom_point() + facet_grid(~  class )
# 

```

## Wrapping a Variable

facet_grid will always create an m x n grid of plots, even if there is no data available for some of the combinations. facet_wrap will only use combinations of variables which have values, so the result isn't always square. facet_wrap will do its best to format the individual plots as neatly as possible. By default, facet_grid will format a single variable as either a single row or single column grid

```{r}

```

## Geoms

Use geom *point* to explicitly specify that a scatterplot should be drawn.

```{r}

```

Next, create the same plot but specify a 'smooth' geometry

```{r}

ggplot(data = mpg, aes(x = displ, y = hwy)) + 
geom_point() + 
geom_smooth()


```

Create the plot again, this time overlay the smooth geometry on the point geometry

```{r}

```

## Task 2 - Creating a Boxplot with qplot

How would you replace the following scatterplot with one containing boxplots?

`ggplot(data = mpg, aex(x = class, y = hwy))`

```{r}
ggplot(data = mpg, aes(x = class, y = hwy)) + geom_boxplot()


```

## Reordering Plots

Reorder the levels of the *class* factor by *hwy*, ascending

```{r}

# this re-orders byt the mean hence not all ascending 
# ggplot(data = mpg, aes(x = reorder(class, hwy), y = hwy)) + geom_boxplot()

# answer - re-order byt he emdian not the mean so all ascending 
ggplot(data = mpg, aes(x = reorder(class, hwy, fun = median),  y = hwy)) + 	geom_boxplot() 


```

## Task 3 - Reordering in Depth

1.  Read the help page for the function `reorder`, here are some tips:
    -   scan page for relevant info.
    -   ignore things that don’t make sense.
    -   try out examples.
2.  Redraw the previous plot with the variable class ordered by the median of hwy.

```{r}

```

### Histograms

Create a histogram for the *carat* feature using default bin-width

```{r}
ggplot(data = diamonds, aes(x = cut, colour = cut)) + ?	geom_bar()


```

Set the binwidth to 1

```{r}

```

Try out a binwidth of 0.01

```{r}

```

## Task 4

Experiment with the bin-width of the *carat* histogram, find a suitable value

```{r}

```

Save the histogram with the suitable bin-width to the current working directory. Call it *diamond_value_by_carat.png*

```{r}

```
