---
title: "R Notebook"
output: html_notebook
---

# Visualizing Data

This notebook can be used to follow along with the practical examples from part 3 of this week's lecture. Before we begin make sure you've installed ggplot2. In these examples we'll be working with the *mpg* dataset. The mpg dataset deals with different makes and model of cars. The **hwy** column gives *highway miles per gallon*. This dataset is used to determine the factors that contribute to a car's miles per gallon

```{r}
# Uncomment the line below to install ggplot2 (if you haven't already).
install.packages('ggplot2')
library(ggplot2)

# The mpg dataset is available by default in R, it has the variable name `mpg`
nrow(mpg)
ncol(mpg)
head(mpg)
tail(mpg)

# count the number of distinct car models in this dataset
length(levels(as.factor(mpg$model)))
```

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point()
```

## Exploring qplot aesthetics

See the lecture notes for details. Plot the *hwy \~ displ* scatterplot above changing each of the 4 aesthetics

```{r}

```

Now setting `class` as the colour attribute

```{r}

```

## Mapping Aesthetics to Values

Use the colour aesthetic to represent each car's *class*

```{r}

```

## Faceting

Use faceting to plot the data in a grid of 1 column and 3 rows. The *drv* feature should determine which row an observation is plotted in

```{r}

```

Now do the same, but rather than rows, have the facets in 3 columns

```{r}

```

## Task 1 - Try each of the 3 examples of faceting

See lecture notes for examples to try

```{r}

```

## Wrapping a Variable

facet_grid will always create an m x n grid of plots, even if there is no data available for some of the combinations. facet_wrap will only use combinations of variables which have values, so the result isn't always square. facet_wrap will do its best to format the individual plots as neatly as possible. By default, facet_grid will format a single variable as either a single row or single column grid

```{r}

```

## Geoms

Use geom *point* to explicitly specify that a scatterplot should be drawn.

```{r}

```

Next, create the same plot but specify a 'smooth' geometry

```{r}

```

Create the plot again, this time overlay the smooth geometry on the point geometry

```{r}

```

## Task 2 - Creating a Boxplot with qplot

How would you replace the following scatterplot with one containing boxplots?

`ggplot(data = mpg, aex(x = class, y = hwy))`

```{r}

```

## Reordering Plots

Reorder the levels of the *class* factor by *hwy*, ascending

```{r}

```

## Task 3 - Reordering in Depth

1.  Read the help page for the function `reorder`, here are some tips:
    -   scan page for relevant info.
    -   ignore things that don’t make sense.
    -   try out examples.
2.  Redraw the previous plot with the variable class ordered by the median of hwy.

```{r}

```

### Histograms

Create a histogram for the *carat* feature using default bin-width

```{r}

```

Set the binwidth to 1

```{r}

```

Try out a binwidth of 0.01

```{r}

```

## Task 4

Experiment with the bin-width of the *carat* histogram, find a suitable value

```{r}

```

Save the histogram with the suitable bin-width to the current working directory. Call it *diamond_value_by_carat.png*

```{r}

```
