---
title: "Converted from IPYNB"
output: html_notebook
---


# Lab Week 3

1.  Using `read.table`, Load the file `shopping.tsv`. into R. `shopping.tsv` is a tab-separated file. Ensure column names are interpreted correctly, and missing values correctly inferred.

```{r}
library(readr)

shopping_tib_1 <- read.table(
  'shopping.tsv',
  sep = '\t',
  header = TRUE,
  stringsAsFactors = TRUE,
  fill = TRUE,
  na.string = c('?', 'Not Applicable')
)
view(shopping_tib_1)

```

2.  Use `readr` to import the table, again ensuring column names and missing values correctly inferred.

```{r}

shopping_tib_1 <- read_tsv(
  'shopping.tsv',
  na= c('?', 'Not Applicable')
)
spec(shopping_tib_1)


```

3.  Import the first 50 rows of `jester-data.csv`. The jester dataset gives ratings given by human raters (columns) to individual jokes (represented by rows). Be aware that this file has no headers. Calculate the average rating of each joke. Tip: check the `t` function.

```{r}

# jester_df1 <- read_csv('jester-data.csv', col_names = FALSE, n_max = 50)
# dim(jester_df1)
# 
# jester_df1$joke_ratingmeans <-(rowMeans(jester_df1, na.rm = TRUE))
# 
# View(jester_df1)
# names(jester_df1)
# dim(jester_df1)
# view(jester_df1)

library(readr)

# Import first 50 rows, no headers
jester_df2 <- read_csv('jester-data.csv', col_names = FALSE, n_max = 50)

# Transpose the data (makes jokes as columns)
# t(jester_df) swaps rows ↔ columns.
#Using t() is especially helpful if you want to calculate means across the transposed dimension without manually thinking in terms of rows or columns.
jester_t2 <- t(jester_df2)

# Compute column means (original rows)
joke_means <- colMeans(jester_t2)

# Add back to original data frame
jester_df2$joke_ratingmeans <- joke_means

# View
View(jester_df2)

# view just the meand column data
jester_df$joke_ratingmeans

#
# Or as a data frame (keeps column name)
jester_df["joke_ratingmeans"]

# View in a separae window
View(jester_df["joke_ratingmeans"])

                               
```

4.  Import the 51st to 100th rows of the jester dataset, and only the first 8 columns. Using R functions, calculate which rater gave the highest overall value of total ratings.

```{r}
jester_tib1 <- read_csv('jester-data.csv',
                        col_names = FALSE,
                        skip = 50,
                        n_max = 50
                        )
# notation is dataframe[ rows, columns]
# before comma which rows to select
# after comma which columns to select
# comma here > select all rows
#
jester_tib2 <- (jester_tib1[,1:8])

View(jester_tib2)

rater_totals <- colSums(jester_tib2, na.rm = TRUE)

# which.max gives col index of the max total
# rater_totals[highest_rater_index] gives the  actual total value
highest_rater_index <- which.max(rater_totals)
highest_rater_total <- rater_totals[highest_rater_index]

highest_rater_index
highest_rater_total 
```

5.  Import the table from the website <https://www.skysports.com/champions-league-table>. Set the correct data type for each column.

```{r}

#install.packages('rvest')
library(rvest)
webpage <- read_html("https://www.skysports.com/champions-league-table")
tableNodes <- html_nodes(webpage, "table")
head(tableNodes)
table <- html_table(tableNodes)
skychamplge_table_df <- table[[1]]
View(skychamplge_table_df)

# gives rows x cols - and tibble or dataframe - tbl_df / tbl / data.frame (tibble is a type of dataframe apparently)
# structure()
str(skychamplge_table_df)
#cass() - gives class of object i.e. tibble or dataframe
class(skychamplge_table_df)

# % > %  means pipe this result into the next function
% > % 
  
# get columns names
colnames(skychamplge_table_df)
# replaces all sequences of whitespace and newlines (\n, \t) with a single space.
# trimws() removes leading and trailing spaces.
clean_names <- trimws(gsub("\\s+", " ", names(skychamplge_table_df)))
clean_names

# get errors when trying to set as.integer etc as column names not clean_names
# so have to clean column names first... as below
names(skychamplge_table_df) <- trimws(gsub("\\s+", " ", names(skychamplge_table_df)))
names(skychamplge_table_df)

# set types for each column
skychamplge_table_df$Position <- as.integer(skychamplge_table_df$Position)
skychamplge_table_df$Team <- as.character(skychamplge_table_df$Team)
skychamplge_table_df$Played <- as.integer(skychamplge_table_df$Played)
skychamplge_table_df$Won <- as.integer(skychamplge_table_df$Won)
skychamplge_table_df$Drawn <- as.integer(skychamplge_table_df$Drawn)
skychamplge_table_df$Lost <- as.integer(skychamplge_table_df$Lost)
skychamplge_table_df$GF <- as.integer(skychamplge_table_df$GF)
skychamplge_table_df$GA <- as.integer(skychamplge_table_df$GA)
skychamplge_table_df$GD <- as.integer(skychamplge_table_df$GD)
skychamplge_table_df$Points <- as.integer(skychamplge_table_df$Points)


```

6.  Save the table as an R object, and then as a comma separated file using the \| character as a colu separator.

```{r}

# use saveRDS for single Object

saveRDS(skychamplge_table_df, "sky_cl_table_rdatobj.rds")

# can reload with
skychamplge_table_df4 <- readRDS("sky_cl_table_rdatobj.rds")
str(skychamplge_table_df4 )

# save a csv using \| as colu separator
save_cvs

# from chat - don't see a save_csv option
# write.table to save a stxt file
# write.csv to save as csv file
# row.names = FALSE ensures R doesn’t add row numbers as an extra column in the CSV.

write.csv(skychamplge_table_df, "cl_csv_save.csv", row.names = FALSE)

# If you want a different separator (like |), you can use write.table() instead:
# quote = FALSE is optional but prevents quotes around text columns
write.table(skychamplge_table_df, "cl_csv_save_table1.csv", sep = '|', row.names = FALSE)

```

7.  Download (using R) the following dataset <https://www3.stats.govt.nz/2018census/Occupied-dwellings-unoccupied-dwellings-dwellings-under-construction-private-and-non-private-dwellings-2006-2013-2018-Censuses-RC-TA-SA2-DHB.zip>. Extract all csv files to separate data frames

```{r}

url <- "https://www3.stats.govt.nz/2018census/Occupied-dwellings-unoccupied-dwellings-dwellings-under-construction-private-and-non-private-dwellings-2006-2013-2018-Censuses-RC-TA-SA2-DHB.zip"
download.file(url = url, destfile = "wk3_zip_dnld.zip")         # Destination file in the current working directory
# the below didn't exrtact into folder
# unzip("wk3_zip_dnld.zip")
# list.files(path = "./wk3_zip_dnld")

# this extracted into folder - craeted it and extracted..then listed the files,,,
unzip("wk3_zip_dnld.zip", exdir = "wk3_zip_dnld")
list.files("wk3_zip_dnld")

# Path to the folder where ZIP was extracted
folder_path <- "wk3_zip_dnld"
(folder_path)

# List all CSV files in the folder ans put the names pf them into a list >> 'csv_files'
csv_files <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)

csv_files

# Read each CSV into a separate data frame in a named list
# lapply() is an R function that applies a function to each element of a list (or vector) and returns a list. Think of it as a “loop in one line.”
# sapply() is very similar to lapply() but with a key difference: it tries to simplify the result.
# ________________________________________
# Comparison
# Function	Returns
# lapply()	Always a list
# sapply()	Vector or matrix if possible, otherwise a list

dfs <- lapply(csv_files, read.csv, stringsAsFactors = FALSE)

# Name each list element after the file (without extension)
names(dfs) <- tools::file_path_sans_ext(basename(csv_files))

# Example: view the first few rows of the first data frame
head(dfs[[1]])
dfs

#dfs is a list of data frames, each corresponding to one CSV file.
# You can access them by name: dfs$FilenameWithoutExtension or by index: dfs[[1]].
# If you want them as separate variables in your environment, you could do

```

8.  Give each data frame a meaningful name. Save the objects to `.rdata` format

```{r}

# list all csv files in he folder_path

folder_path <- "wk3_zip_dnld"

# List CSV files
csv_files <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE)
csv_files

# Create clean variable names
df_names <- tools::file_path_sans_ext(basename(csv_files))
df_names <- gsub("-", "_", df_names)        # replace hyphens with underscores
df_names <- gsub(" ", "_", df_names)        # replace spaces with underscores
df_names <- make.names(df_names)            # ensure valid R variable names

# Read CSV files into a list
dfs <- lapply(csv_files, read.csv, stringsAsFactors = FALSE)

# Assign names to the list elements
names(dfs) <- df_names


list2env(dfs, envir = .GlobalEnv)

save(list = df_names, file = "wk3_dataframes.rdata")

load("wk3_dataframes.rdata")



```
